# Peer-Assessment Assignment 1
                       
## Loading and Processing the data
```{r loading data, echo=TRUE}
setInternet2(TRUE)
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp) 
data <- read.csv(unz(temp, "activity.csv"))
data$date <- as.Date(data$date)
unlink(temp)
```



## Mean total number of steps taken per day
### Summary for total number of steps taken each day
```{r, echo = TRUE}
DailyTotalSteps <- aggregate(data$steps, list(data$date), sum)
colnames(DailyTotalSteps) <- c("date", "steps")
```

### Histogram of the total number of steps taken each day
```{r, echo = TRUE}
barplot(DailyTotalSteps$steps, names.arg = DailyTotalSteps$date, 
    xlab = "Date", ylab = "Steps",
    main = "Total Number of Steps Taken Each Day", axis.lty = 1)
```

### Mean and median of step taken each day
```{r, echo = FALSE}
MeanValue <- mean(DailyTotalSteps$steps, na.rm = TRUE)
MedianValue <- median(DailyTotalSteps$steps, na.rm = TRUE)
```
Mean and median total number of steps taken per day: `r as.integer(MeanValue)` and `r MedianValue`



## Average daily activity pattern
### Time series plot
```{r, echo = TRUE}
AverageSteps <- aggregate(steps~interval, data = data, FUN = "mean")
plot(AverageSteps$interval, AverageSteps$steps, type = "l",
    xlab = "5-minute interval", ylab = "Average nubmer of steps taken",
    main = "Average number of steps taken by 5-minute interval")
```

### 5-minute interval with maximum number of steps
```{r, echo = FALSE}
Interval_maxvalue <- AverageSteps$interval[AverageSteps$steps == max(AverageSteps$steps)]
```
5-minute interval with maximum number of steps of average daily activity: `r Interval_maxvalue`



## Imputing missing values
### Total number of missing values in the dataset
```{r, echo = FALSE}
MissingCount <- sum(is.na(data$steps))
```
Total number of missing value in the dataset: `r MissingCount`

### Imputing missing values with the mean for the 5-minute interval
```{r, echo = FALSE}
colnames(AverageSteps)[2] <- "AverageSteps"
ImputData <- merge(data, AverageSteps, by.x = "interval", by.y = "interval")
ImputData$steps[is.na(ImputData$steps)] <- ImputData$AverageSteps[is.na(ImputData$steps)]
```

### Histogram with imputing data
```{r, echo = TRUE}
ImputTotalSteps <- aggregate(steps~date, data = ImputData, FUN = "sum")
barplot(ImputTotalSteps$steps, names.arg = ImputTotalSteps$date, 
    xlab = "Date", ylab = "Steps",
    main = "Total Number of Steps Taken Each Day after imputing", 
    axis.lty = 1)
```

### Mean and median of step taken each day with imputing data
```{r, echo = FALSE}
ImputMeanValue <- mean(ImputTotalSteps$steps, na.rm = TRUE)
ImputMedianValue <- median(ImputTotalSteps$steps, na.rm = TRUE)
if(MeanValue > ImputMeanValue) {
    different1 <- "Increase"
    different2 <- "Yes"
} else if(MeanValue <ImputMeanValue) {
    different1 <- "Decrease"
    different2 <- "Yes"
} else {
    different1 <- "Equal"
    different2 <- "No"
}
if(MedianValue > ImputMedianValue) {
    different3 <- "Increase"
    different4 <- "Yes"
} else if(MedianValue <ImputMedianValue) {
    different3 <- "Decrease"
    different4 <- "Yes"
} else {
    different3 <- "Equal"
    different4 <- "No"
}

```
Mean and median total number of steps taken per day: `r as.integer(ImputMeanValue)` and `r as.integer(ImputMedianValue)`

Do these values different from the previous estimates: `r different2`

What is the impact of imputing: 
Mean Value `r different1` from `r as.integer(MeanValue)` to `r as.integer(ImputMeanValue)`
Median Value `r different3` from `r MedianValue` to `r as.integer(ImputMedianValue)`

## Differences in activity patterns between weekdays and weekends
```{r, echo = TRUE}
ImputData$weekday <- weekdays(ImputData$date) 
ImputData$weekday[ImputData$weekday %in% c("Saturday", "Sunday")] <- "Weekend"
ImputData$weekday[ImputData$weekday != "Weekend"] <- "Weekday"
ImputData$weekday <- as.factor(ImputData$weekday)
WeekdayData <- aggregate(steps~interval + weekday, data = ImputData, FUN = "mean")
WeekdayData$steps <- format(WeekdayData$steps, scientific = FALSE) 
library(lattice)
xyplot(as.integer(steps) ~ interval | weekday, data = WeekdayData, layout = c(1, 2),
       type = "l", xlab = "Interval", ylab = "Number of steps")

```
It is different in activity patterns between weekdays and weekends.

